---
title: "Your Project Title (be descriptive and creative!)"
subtitle: "You can either put a subtitle here or delete this line"
author: "Team members' full names, tutorial groups, and group number"
format:
  revealjs:
    theme: simple
    slide-number: true
    fontsize: 3.5em
    width: 1600
    height: 900
    margin: 0.1
    pdf-max-pages-per-slide: 1
    self-contained: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,      # Hide code   
  warning = FALSE,   # Suppress warnings
  message = FALSE   # Suppress messages
  
)

# Note: the options above are to hide the code chunk in the slides and to not include the code in the slides
# There are many other options you can use to customize which parts of the code are run / displayed, 
# as well as how the output is displayed. 
# You can learn more at https://quarto.org/docs/computations/execution-options.html

```

```{r load-data}

# setup
library(tidyverse)

# Load data we need to load the data, but don't want to display it in our slides
# The code chunk options at in the curly brackets above specif
data_patient_members <- read_csv("Member_Background.csv")
data_attendance_records <- read_csv("Service_Deliveries.csv")


## Parse members data
## Types of cancer - convert to set of indicator variables for different types of cancer (don't include ones which are too rare)
data_patient_members <- data_patient_members |>
   mutate(lung = str_detect(types_of_cancer, "Lung"),
         breast = str_detect(types_of_cancer, "Breast"),
         brain = str_detect(types_of_cancer, "Brain"),
         leukemia = str_detect(types_of_cancer, "Leukemia"),
         prostate = str_detect(types_of_cancer, "Prostate"),
         colorectal = str_detect(types_of_cancer, "Colorectal"),
         pancreatic = str_detect(types_of_cancer, "Pancreatic"),
         lymphoma_non_hodgkins = str_detect(types_of_cancer, "Lymphoma, Non-hodgkins"),
         liver = str_detect(types_of_cancer, "Liver"),
         multiple_myeloma = str_detect(types_of_cancer, "Multiple Myeloma"),
         bladder = str_detect(types_of_cancer, "Bladder"),
         cervical = str_detect(types_of_cancer, "Cervical"),
         esophagus = str_detect(types_of_cancer, "Esophagus"),
         gallbladder_and_bile_duct = str_detect(types_of_cancer, "Gallbladder and Bile Duct"),
         head_and_neck = str_detect(types_of_cancer, "Head and Neck"),
         kidney = str_detect(types_of_cancer, "Kidney"),
         lymphoma_hodgkins = str_detect(types_of_cancer, "Lymphoma, Hodgkins"),
         ovarian = str_detect(types_of_cancer, "Ovarian"),
         skin = str_detect(types_of_cancer, "Skin"),
         stomach = str_detect(types_of_cancer, "Stomach"),
         testicular = str_detect(types_of_cancer, "Testicular"),
         thyroid = str_detect(types_of_cancer, "Thyroid"),
         cervical = str_detect(types_of_cancer, "Cervical"),
         other = str_detect(types_of_cancer, "Other"))


## Impairments - convert to set of indicator variables for different types of impairment (don't include ones which are too rare)
data_patient_members <- data_patient_members |>
  mutate(visual = str_detect(i_have_the_following_impairment, "visual"),
         cognitive = str_detect(i_have_the_following_impairment, "cognitive"),
         limited_mobility = str_detect(i_have_the_following_impairment, "mobility"),
         none_of_the_above = str_detect(i_have_the_following_impairment, "None of the above"),
         other = str_detect(i_have_the_following_impairment, "other"),
         speech = str_detect(i_have_the_following_impairment, "speech"),
         hearing = str_detect(i_have_the_following_impairment, "hearing")) 

## Program interests: Split into boolean indicators for different categories
data_patient_members <- data_patient_members |>
  mutate(therapeutic_arts = str_detect(program_interest, "Therapeutic Arts"),
         educational_and_self_developoment = str_detect(program_interest, "Educational & Self-Development"),
         symptom_management = str_detect(program_interest, "Symptom Management"),
         social_events = str_detect(program_interest, "Social / Events"),
         financiaL_and_workplace_strategies = str_detect(program_interest, "Financial & Workplace Strategies"),
         exercise_and_movement = str_detect(program_interest, "Exercise & Movement"),
         individual_and_group_support = str_detect(program_interest, "Individual & Group Support")) |>
  mutate(num_program_interests = therapeutic_arts + educational_and_self_developoment + symptom_management + social_events + financiaL_and_workplace_strategies + exercise_and_movement + individual_and_group_support)

## How did you hear about wellspring: Split into booleans
data_patient_members <- data_patient_members |>
  mutate(word_of_mouth = str_detect(how_have_you_heard_of_wellspring, "Word of mouth"),
         flyer = str_detect(how_have_you_heard_of_wellspring, "flyer"),
         wellspring_member = str_detect(how_have_you_heard_of_wellspring, "Member"),
         wellspring_member = str_detect(how_have_you_heard_of_wellspring, "Volunteer"),
         doctor = str_detect(how_have_you_heard_of_wellspring, "Doctor"),
         hospital = str_detect(how_have_you_heard_of_wellspring, "hospital"),
         hospital = str_detect(how_have_you_heard_of_wellspring, "healthcare"),
         tv = str_detect(how_have_you_heard_of_wellspring, "television"),
         website = str_detect(how_have_you_heard_of_wellspring, "Website"),
         family = str_detect(how_have_you_heard_of_wellspring, "Family"),
         social_media = str_detect(how_have_you_heard_of_wellspring, "Social Media"),
         other = str_detect(how_have_you_heard_of_wellspring, "Other"))

```

```{r}
# dataset dimensions
data_attendance_records |> 
  dim()

# What members register for the most events?
data_attendance_records |> 
  group_by(member_id) |>
  summarise(n=n()) |>
  arrange(desc(n))

# Let's examine information on the member with the highest attendance
high_attendance <- data_attendance_records |> 
  filter(member_id == "003ON000007sHYtYAM")

data_patient_members |> filter(member_id == "003ON000007sHYtYAM")

# What is range of registration dates for members in first table
data_patient_members |> 
  group_by(member_start_year) |>
  summarise(n=n()) |>
  arrange(desc(n))
```

## A heading that starts on a new page

This document shows a few basics on making slides with R markdown.

To produce the slides, Render this qmd document by clicking at the top of the screen

You will need to submit a qmd and pdf version of your slides.

You must use the same slides in your video presentation and at the in-person poster fair (printed slides)

## Small font size {.smaller}

You can also start a new slide with a header (indicated by two hashtags)

This slide will have smaller font size because of what we specified right after the header

## Mixed font size on a single slide

::: {style="font-size: 0.8em;"}
This div has a reduced font size through inline styling.
:::

::: {style="font-size: 4em;"}
And now it's bigger
:::

## A slide with a bulleted list

-   Bullet 1
-   Bullet 2

## A slide with a numbered list

1.  First point
2.  Second point

# New Section by using a bigger header (one \# sign)

## First slide in new section - multiple columns

::: columns
::: {.column width="40%"}
-   Bullet 1 in left
-   Bullet 2 in left
:::

::: {.column width="60%"}
Right column
:::
:::

# Research Question 3

## Statement of the Question

**Does the likelihood of missing events differ between in-person and online programs?** This question explores whether the format of a program (in-person vs. online) influences attendance rates. Answering this can help Wellspring understand engagement levels.

------------------------------------------------------------------------

## Relevant Visualization

```{r}
df <- left_join(data_patient_members, data_attendance_records, by = "member_id")

df_clean <- df |> filter(!is.na(service_session_location), !is.na(attendance_status)) |> mutate(missed = attendance_status == "Unexcused Absence", event_format = ifelse(str_detect(service_session_location, "Online"), "Online", "In-Person"))

missed_summary <- df_clean |> group_by(event_format) |>
summarise(n_total = n(), n_missed = sum(missed),prop_missed = n_missed / n_total)

missed_plot <- ggplot(missed_summary, aes(x = event_format, y = prop_missed, fill = event_format)) + geom_bar(stat = "identity") + labs(title = "Proportion of Missed Events: Online vs In-Person" , x = "Event Format", y = "Proportion Missed") + theme_minimal()

missed_plot

```

## Set-up of Statistical Model/Method

We use a two-sample hypothesis test for proportions.

Let p1 = proportion of missed in-person events\
Let p2 = proportion of missed online events

H₀: p1 = p2\
Hₐ: p1 ≠ p2

We simulate the difference in missed proportions by randomly shuffling the event format labels.

## Hypothesis Test

```{r}
obs_diff <- df_clean |> group_by(event_format) |> summarise(prop_missed = mean(missed)) |> summarise(diff = diff(prop_missed)) |> pull(diff)

set.seed(130)
nsim <- 1000
simulated_diffs <- rep(NA, nsim)

for (i in 1:nsim) {shuffled <- df_clean |> mutate(shuffled_loc = sample(event_format)) |>
group_by(shuffled_loc) |> summarise(prop_missed = mean(missed)) |> summarise(diff = diff(prop_missed)) |> pull(diff)

simulated_diffs[i] <- shuffled
}

null_dist <- ggplot(tibble(simulated_diffs), aes(x = simulated_diffs)) + geom_histogram(bins = 30, fill = "gray", color = "black") + geom_vline(xintercept = obs_diff, color = "red") + geom_vline(xintercept = -obs_diff, color = "red") + labs(title = "Simulated Null Distribution of Difference in Missed Proportions", x = "Difference (In-person - Online)", y = "Frequency"
  )

null_dist

```

## Results and Interpretation

```{r}
p_value <- (sum(abs(simulated_diffs) >= abs(obs_diff)) + 1) / (nsim + 1)
p_value

```

The p-value is approximately 0.001, meaning that in only 0.1% of simulated random assignments did we observe a difference in missed proportions as extreme as the one in our data. This is strong evidence against the null hypothesis. Thus, I can conclude that the proportion of missed events is at the very least different between in-person and online formats. This suggests that event format does affect attendance behavior and may require different engagement strategies.
